### 2024-10-22 paper/paper.qmd
Updated paper with data and model section

---
title: "An abundance of invasive trees in Toronto and a lack of trees in the streets of Toronto's core. "
subtitle: ""
author: 
  - Veyasan Ragulan
thanks: "Code and data are available at: https://github.com/Veyasan1/2024USElectionModel"
date: today
date-format: long
abstract: "Tree- lined streets are an exceptional way to add nature to urban envirnoments. Using the Street Tree data from OpenData Toronto, this paper examines the patterns and trends in Toronto's street trees. Tree species count, count of trees by ward, and distribution of tree-diameters were analysed. Uncovered was a abundance of a non-native tree species, and a lack of trees on the streets of Toronto's core"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

In this paper we look at polling data from 538 on the 2024 US Presidential election, specifically on Demoncratic Candidate Kamala Harris. We will construct a model that will be used to make a prediction on whether Harris wins the US Presidency

Section 2 will outline the source of this data. Section 3 covers the model and it's parameters. Section 4 is where discussion will be made about the models predictions and how realistically they line up with current affairs. Finally, section 4 discusses any weakness and limitations that can be considered for another report.

# Data {#sec-data}

The data used in this model comes from fivethirtyeight.com, a polling site known for their in-depth analysis and data on US elections. Here, we have used the 2024 general election data, originally accessed on 10/21/2024. The data contains information from several polls taken across the United States, including the operator, sponsors, methodology, a numeric grade to quanitfy the quality of the poll, and when and where the poll was taken. From this dataset, polls predicting a Harris win were considered, to construct the predictive model.

# Model

The model contains 4 parameters. They are pollster, state, sponsors, and methodology.

Pollster is the operator of the observed poll. Pollster provides a relatively easy way to identify trends in certain polls, depending on who is conducting them.

State tells us where the poll was conducted. Across the United States, ther are states that traditionally vote Democrat (New York, California), some that traditionally vote Republican (Texas, Florida), and 'swing states' (Michigan, Pennsylvania). Swing states in particular are the most interesting states to watch as votes from these states tend to win a candiate the US presidency.Knowing where the polls have been taken also lets us understand if their choice was one bound by the state's culture.

Sponsors are organizations that fund or otherwise assist the pollster in operating the poll. Sponsors carry the risk of bias with them, potentially influencing the pollster, and thus the polls to skew in a direction the sponsor favors preferably. Knowing who these sponsors are potemntially lets us make informed predictions on the outcome of polls.

Methodology is the way the poll was conducted. Sometimes the composition of the poll influences peoples votes, so this is an interesting parameter to track. This also lets us see if certain methods yield better or worse predictions.

# Results

```{r}
#| echo: false
#| warning: false
#| message: false

# Posterior predictive checks
pp_check(bayesian_model)

# Summarize the model
summary(bayesian_model)

# Plot random effects
plot(bayesian_model, pars = "(Intercept)", prob = 0.95)

```

```{r}
#| label: fig-1
#| fig-cap: Street tree count of 10 most popoulus tree species in Toronto up to September 2024
#| echo: false
#| warning: false
#| message: false

# Summarize the model
summary(spline_model)

# Posterior predictive checks
pp_check(spline_model)
```

\newpage

# Discussion

## The worrying prominence of Acer platanoides

## The lack of trees in Toronto-Centre

## Anomalies in DBH diameters and potential causes

## Weaknesses and next steps

\newpage

# Appendix

## Data Cleaning

# References

### 2024-10-22 paper/paper.qmd
Updated README
Exported models to models subfolder

Imported models to Quarto doc

Updated references and general structure

Small edits to clean_data

Deleted unnecessary files

@@
 ---
-title: "An abundance of invasive trees in Toronto and a lack of trees in the streets of Toronto's core. "
+title: "A look into Kamala Harris' polls in the upcoming 2024 US Presidential Election "
 subtitle: ""
@@
 date-format: long
-abstract: "Tree- lined streets are an exceptional way to add nature to urban envirnoments.
-Using the Street Tree data from OpenData Toronto, this paper examines the patterns and trends in Toronto's street trees.
-Tree species count, count of trees by ward, and distribution of tree-diameters were analysed.
-Uncovered was a abundance of a non-native tree species, and a lack of trees on the streets of Toronto's core"
+abstract: "The 2024 US election has been seen as a close race between the Democratic nominess Kamala Harris, and the Republican Nominee Donald Trump, with each candidate holding down key states and matching each other in battleground states such as Pennsylvania and Georgia.
+This paper looks at the polling as of late October 2024 to see how Kamala Harris fares as the election draws near."
 format: pdf
@@
 library(tidyverse)
+
+library(rstanarm)
+
+bayesian_model <-
+  readRDS(file = here::here("models/electionmodel.rds"))
+
+spline_model <- 
+  readRDS(file = here::here("models/splinemodel.rds"))
 ```
@@
 
+This report was written with the assistance of R @citeR.
+Additionally, @tidyverse, and @rstanarm were used to clean the data, write models, and analyse results.
+The structure of this report heeds Rohan Alexander's example @tellingstories.
+
 # Data {#sec-data}
 
-The data used in this model comes from fivethirtyeight.com, a polling site known for their in-depth analysis and data on US elections.
+The data used in this model comes from fivethirtyeight.com, a polling site known for their in-depth analysis and data on US elections @fivethirtyeight.
 Here, we have used the 2024 general election data, originally accessed on 10/21/2024.
@@
 
-## The worrying prominence of Acer platanoides
+## Trends from pollsters
 
-## The lack of trees in Toronto-Centre
+## Trends from states
 
-## Anomalies in DBH diameters and potential causes
+## Trends from sponsors
+
+## Trends from methodology
 
@@
 
-## Data Cleaning
+## Pollster Analysis
+
+## Poll Blueprint
 

### 2024-10-22 scripts/03-clean_data.R

@@ -4,7 +4,7 @@
 # Date: 18 October 2023
 # Contact: veyasan.ragulan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: president_polls.csv, which is the presidental general election poll sourced from 538
+# Pre-requisites: president_polls.csv, sourced from https://projects.fivethirtyeight.com/polls/ (select Presidential general election polls current cycle)
 
 #### Workspace setup ####
 library(tidyverse)
@@ -24,7 +24,7 @@ data <- read_csv("data/01-raw_data/president_polls.csv") |>
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+    numeric_grade >= 3 # Need to investigate this choice - come back and fix. 
     # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
   ) |>
   mutate(

### 2024-10-22 scripts/06-model_data.R

@@ -82,3 +82,14 @@ summary(spline_model)
 # Posterior predictive checks
 pp_check(spline_model)
 
+# Save model
+saveRDS(
+  bayesian_model,
+  file = "models/electionmodel.rds"
+)
+
+saveRDS(
+  spline_model,
+  file = "models/splinemodel.rds"
+)
+

### 2024-10-24 scripts/04-test_analysis_data.R
Edited test-analysis. Added code for parquet

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests both simulation and real data used in analysis
+# Author: Veyasan Ragulan
+# Date: 24 September 2024
+# Contact: veyasan.ragulan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: A simulation dataset, and the real dataset, obatined from https://projects.fivethirtyeight.com/polls/
+# Any other information needed? 
 
 
 #### Workspace setup ####

### 2024-10-24 scripts/03-clean_data.R
Added edits to write a parquet file

@@ -14,6 +14,7 @@ library(broom)
 library(modelsummary)
 library(rstanarm)
 library(splines)
+library(arrow)
 
 #### Clean data ####
 # Read in the data and clean variable names
@@ -38,3 +39,6 @@ just_harris_high_quality <- data |>
 
 #### Save data ####
 write_csv(just_harris_high_quality, "data/02-analysis_data/harris.csv")
+
+write_parquet(just_harris_high_quality, "data/02-analysis_data/harris.parquet")
+

### 2024-11-03 paper/paper.qmd
Update paper.qmd
Updated the quarto document to reflect changes made offline. Changes include a rewrite of the introduction, now only needing results, updated a bit of the data section, and inserted the methodology section to the report

@@
 
-In this paper we look at polling data from 538 on the 2024 US Presidential election, specifically on Demoncratic Candidate Kamala Harris.
-We will construct a model that will be used to make a prediction on whether Harris wins the US Presidency
+The 2024 US Presidential race has been closely contested since incumbent president, Joe Biden, announced his exit from the race.
+The race is primarily between current VP, Democratic Party candidate Kamala Harris (with running mate Time Walz), and former president, Republican Party candidate Donald Trump (with running mate J.D.
+Vance).
+Polls across the country have conflicting reports on who will come out on top, with some sources suggesting an unheard-of tie in the electoral college vote (269-269 vote distribution) [@tieVote].
+
+To get a better grasp of nationwide and state opinons going into the election, a predictive model was created in this report.
+The model utilises data from fivethirtyeight.com, which hosts a dataset of diverse and reputable polls taken across the country.
+Using this data, we (INSERT WRANGLING DONE TO DATASHEET).
+Then we constructed a generalized linear model, predicting (INSERT PREDICTOR), using (INSERT PREDICTEES) as predictors.
+The results from the model show (INSERT RESULTS FROM MODEL)
 
@@
 
-This report was written with the assistance of R @citeR.
-Additionally, @tidyverse, and @rstanarm were used to clean the data, write models, and analyse results.
-The structure of this report heeds Rohan Alexander's example @tellingstories.
-
 # Data {#sec-data}
 
-The data used in this model comes from fivethirtyeight.com, a polling site known for their in-depth analysis and data on US elections @fivethirtyeight.
-Here, we have used the 2024 general election data, originally accessed on 10/21/2024.
-The data contains information from several polls taken across the United States, including the operator, sponsors, methodology, a numeric grade to quanitfy the quality of the poll, and when and where the poll was taken.
-From this dataset, polls predicting a Harris win were considered, to construct the predictive model.
+The dataset provided by fivethirtyeight contains over 15000 observations across 50 variables. Each observation is an individual poll done on the topic of the 2024 US presidency.
+
+# Methodology
+
+The dataset comes from fivethirtyeight.com, a polling analysis website originally run by Nate Silvers, who is no longer affiliated with the website.
+Fivethirtyeight takes polls conducted on a particular topic (in this case the 2024 US presidency), and aggregates them in one dataset.
+Given the diverse array of polls conducted, there is the possiblity of misleading polls and faulty data making its way into the aggregated dataset.
+This impacts the usefulness of models trained on this data, and any analysis that comes out of either the model, the dataset, or both.
+To combat this issue, fivethirtyeight implements its own quality control on the data being aggregated.
+The following breakdown of checks fivethirtyeight uses comes courtesy of their FAQ and methodology article, which is up to date as of September 24 2024 [@faq538].
+
+Fivethirtyeight routinely checks up on pollsters as they produce reports, wther they have an established track record, or if they are just new to data presentation, There are 2 types of checks fivethirtyeight uses: Methodology and Ethics.
+
+Methodology refers to the scientific rigor exhibited by the pollster, when conducting the poll, and in the presentation of the results.
+Fivethirtyeight requires each poll to have or easily obtain the following:
+
+-   Pollster Identity
+
+-   Survey Dates
+
+-   Population Sample (Size and other distinct attributes)
+
+-   Polling Method (How was the poll conducted?)
+
+-   Sponsors (Identity and Amount
+
+Even if these conditions are met, fivethirtyeight may refuse to include certain polls in their data, due to flaws in methodology. Some examples include:
+
+-   Polls with inappropriate sample for target population
+
+-   Polls based upon predictive models
+
+-   Polls that implement recontact with participants (potential for bias to creep into the sample)
+
+-   Polls done by amateur/non-professional pollsters
+
+-   Polls with an "informed ballot" (when information about a candidate is given to the participant before they are asked who they would vote for)
+
+The second check fivethirtyeight implements is ethical standard, adapted from the American Association for Public Opinion Research's Code of Professional Ethics and Practices .
+Fivethirtyeight may refuse to include polls that:
+
+-   Fabricate or falsify data
+
+-   Are associated with the betting industry
+
+-   Refuse to disclose their methodology, either in their findings, or after being contacted for clarification
+
+-   Misrepresent their true purpose (being part of campaign analysis or done for a particular party or candidate)
+
+-   Disclose errors that come about from their work, and rectify them as best as possible
+
+-   Utilize methods that will give misleading results
+
+Fivethirtyeight will conduct an initial check on pollsters, seeing if their methodology and ethics are publicly available, and are in agreement with fivethirtyeight's checks.
+If unclear, fivethirtyeight will do a more thorough investigation of the pollster, and ask the organizers questions about their methodology and ethics if none were visible in the public report.
+If there are any violations in ethics, the severity on consequences will vary on wjat the offence is.
+If there is evidence of falifyin data or engaging in betting, the pollster is blacklisted from their data, any polls that come out of this pollster are extremely dubious and not trustworthy to build a model or report on.
+Otherwise, pollsters are given a chance to fix ethical issues presented by fivethirtyeight.
+Failiure to do so results in the pollster and any of its polls being taken off the dataset, but if the pollster shows it has updated its ethics policies and applied them to recent reports, they may be allowed back in.
+
+Partisan polls are a small subset of polls that are made with the backing or organizations affiliated with one or more political organizations.
+Fivethirtyeight will allow these polls into their data, but they will be marked as "Partisan", and extra checks are put in place to ensure the polling data hasn't been influenced unethically by the pollster or it's affiliated organizations.
 

### 2024-11-03 paper/paper.qmd
Updated Data and references files

@@
 ```{r}
-#| include: false
+#| include: true
 #| warning: false
@@
 
-# Methodology
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+library(tidyverse)
+
+library(rstanarm)
+
+just_harris_high_quality <-
+  read_csv(file = here::here("data/02-analysis_data/harris.csv"))
+
+#### Plot data ####
+base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# Color by pollster
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+# Facet by pollster
+# Make the line 21 issue obvious
+# Also - is there duplication???? Need to go back and check
+base_plot +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(vars(pollster))
+
+# Color by pollscore
+base_plot +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+```
+
+# Measurement
 

### 2024-11-03 paper/paper.qmd
Updated Measurement section, added sources to references.bib, updated raw data presidential_polls.csv, and re-rendered the quarto document

@@
 ```{r}
-#| include: true
+#| include: false
 #| warning: false
@@
 library(rstanarm)
+
+just_harris_high_quality <-
+  read_csv(file = here::here("data/02-analysis_data/harris.csv"))
 
@@
 Vance).
-Polls across the country have conflicting reports on who will come out on top, with some sources suggesting an unheard-of tie in the electoral college vote (269-269 vote distribution) [@tieVote].
+Polls across the country have conflicting reports on who will come out on top, with some sources suggesting an unheard-of tie in the eletoral college vote (269-269 vote distribution) [@tieVote].
 
-To get a better grasp of nationwide and state opinons going into the election, a predictive model was created in this report.
+To get a better grasp of nationwide and state opinions going into the election, a predictive model was created in this report.
 The model utilises data from fivethirtyeight.com, which hosts a dataset of diverse and reputable polls taken across the country.
@@
 
+### Key Variables
+
+#### **`transparency_score:`**
+
+#### **`poll_score:`**
+
 ```{r}
-#| include: false
+#| include: true
 #| warning: false
@@
 
-library(tidyverse)
+#### Plot data ####
+base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
 
-library(rstanarm)
+# Color by pollscore
+base_plot +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
 
-just_harris_high_quality <-
-  read_csv(file = here::here("data/02-analysis_data/harris.csv"))
+```
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
 
@@
 
-The dataset comes from fivethirtyeight.com, a polling analysis website originally run by Nate Silvers, who is no longer affiliated with the website.
+The dataset comes from fivethirtyeight.com, a statistical analysis website dedicated to US sports and elections [@fivethrityeight].
 Fivethirtyeight takes polls conducted on a particular topic (in this case the 2024 US presidency), and aggregates them in one dataset.
-Given the diverse array of polls conducted, there is the possiblity of misleading polls and faulty data making its way into the aggregated dataset.
-This impacts the usefulness of models trained on this data, and any analysis that comes out of either the model, the dataset, or both.
-To combat this issue, fivethirtyeight implements its own quality control on the data being aggregated.
-The following breakdown of checks fivethirtyeight uses comes courtesy of their FAQ and methodology article, which is up to date as of September 24 2024 [@faq538].
 
-Fivethirtyeight routinely checks up on pollsters as they produce reports, wther they have an established track record, or if they are just new to data presentation, There are 2 types of checks fivethirtyeight uses: Methodology and Ethics.
+Election polls are used to gauge pulic opinion about candidates, political topics, voter angagement, and public opinion of the election process.
+Election polls also serve as key insight to political organisations representing candidates, polls can show potenital weak points are low support in areas and populations.
+
+Election polls are similarly to regular polls, they involve asking a sample population a series of questions via a delivery method.
+The key distinction is that election polls typically try to target *voters*, or people who are most likely to vote in the election cycle being investigated.
+This is a key challenge to pollsters, as there are many subgroups of voters, such as people who seldom vote or almost never vote.
+According to Pew Research, a third of US citizens did not vote in the 2020 election, despite that election cycle having high voter turnout [@pew_polls].
+There may also be cases when repondents of a poll saying they will/will not vote change their mind come election day.
+
+Election polling has exploded in recent times according to Pew Research.
+They suggest that double thre are double the amount of active pollsters in 2022 compared to 2000 [@pew_polls].
+One reason mentioned is the advent of online polling, which is simple, cheap, and can easily reach US citizens across the country.
+As a result, while the diversity of pollsters has increased dramatically, this also means there are as many amateur polls being conducted.
+
+Errors in election polling have been an issue in the last two cycles of presidential elections.
+Between the 2016 and 2020 elections, there has been a trend of pollsters not fully capturing Trump's support.
+For example in 2020, despite Biden winning the election, there were reports from the American Associcate for Public Opinion Reseach that national polls overstated his victory over Trump by 3.9 percentage points [@pew_polls].
+
+Fivethirty eight mitigates the influx of pollsters and the potential for errors by routinely checking up on pollsters as they produce reports, wther they have an established track record, or if they are just new to data presentation, There are 2 types of checks fivethirtyeight uses: Methodology and Ethics.
 
 Methodology refers to the scientific rigor exhibited by the pollster, when conducting the poll, and in the presentation of the results.
-Fivethirtyeight requires each poll to have or easily obtain the following:
+Fivethirtyeight requires each poll to have or easily obtain the following [@faq538]:
 
@@
 
-Even if these conditions are met, fivethirtyeight may refuse to include certain polls in their data, due to flaws in methodology. Some examples include:
+Even if these conditions are met, fivethirtyeight may refuse to include certain polls in their data, due to flaws in methodology. Some examples include [@faq538]:
 
@@
 The second check fivethirtyeight implements is ethical standard, adapted from the American Association for Public Opinion Research's Code of Professional Ethics and Practices .
-Fivethirtyeight may refuse to include polls that:
+Fivethirtyeight may refuse to include polls that [@faq538]:
 
@@
 Otherwise, pollsters are given a chance to fix ethical issues presented by fivethirtyeight.
-Failiure to do so results in the pollster and any of its polls being taken off the dataset, but if the pollster shows it has updated its ethics policies and applied them to recent reports, they may be allowed back in.
+Failiure to do so results in the pollster and any of its polls being taken off the dataset, but if the pollster shows it has updated its ethics policies and applied them to recent reports, they may be allowed back in [@faq538].
 
 Partisan polls are a small subset of polls that are made with the backing or organizations affiliated with one or more political organizations.
-Fivethirtyeight will allow these polls into their data, but they will be marked as "Partisan", and extra checks are put in place to ensure the polling data hasn't been influenced unethically by the pollster or it's affiliated organizations.
+Fivethirtyeight will allow these polls into their data, but they will be marked as "Partisan", and extra checks are put in place to ensure the polling data hasn't been influenced unethically by the pollster or it's affiliated organizations [@faq538].
 

### 2024-11-04 paper/paper.qmd
Added kableExtra package, used it to generate summary tables in Data section of qmd

@@
 library(tidyverse)
-
+library(kableExtra)
 library(rstanarm)
@@
 
-The dataset provided by fivethirtyeight contains over 15000 observations across 50 variables. Each observation is an individual poll done on the topic of the 2024 US presidency.
+The dataset provided by fivethirtyeight contains over 15000 observations across 50 variables.
+Each observation is a poll conducted on the 2024 US Presidential Election.
+Pollsters are required by fivethirtyeight to disclose certain information about their poll, either publicly, or through an investigation.
+Of this information, a handful are key to the predictive model.
 
-### Key Variables
+## Predictors
 
-#### **`transparency_score:`**
+### **`poll_score:`**
 
-#### **`poll_score:`**
+The error and bias associated with a pollster.
+Positive values indicate more bias and errors, negative values indicate less.
+This value will tell us if biases in polls influences the predicted percentages of each candidate
 
@@
   theme(legend.position = "bottom")
+
+pollscore_summary <- just_harris_high_quality %>%
+  summarize(
+    Min = min(pollscore, na.rm = TRUE),
+    Mean = mean(pollscore, na.rm = TRUE),
+    Median = median(pollscore, na.rm = TRUE),
+    Max = max(pollscore, na.rm = TRUE)
+  )
+
+pollscore_summary %>% 
+  kable() %>%
+  kable_styling()
+
+```
+
+### `sample_size`
+
+Polls are conducted over a subset of the targeted population (Hispanics, young people, Texans, LGBTQ+, rural communities, etc.).
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+
+#### Plot data ####
+base_plot +
+  geom_point(aes(color = factor(sample_size))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+```
+
+### `state`
+
+The state that each poll as conducted in. If there are any empty values, we assume these are national polls.
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+
+#### Plot data ####
+just_harris_high_quality %>%
+  ggplot(mapping=aes(x = state), y=count(n)) +
+  geom_bar()
+
+state_summary <- just_harris_high_quality %>%
+  summarize(
+    Min = min(count(state), na.rm = TRUE),
+    Max = max(state, na.rm = TRUE)
+  )
+
+state_summary %>% 
+  kable() %>%
+  kable_styling()
 
@@
 
+### `party_binary`
+
+### `end_date`
+
+Every poll has a start and end date for querying its participants.
+Polls that are conducted closer to election day may be more reflective of American opinions as opposed to polls conducted in August or September.
+This is why we will be using this variable as one of our predictors in the model.
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+
+#### Plot data ####
+just_harris_high_quality %>%
+  ggplot(mapping=aes(x = end_date)) +
+  geom_histogram()
+
+end_date_summary <- just_harris_high_quality %>%
+  summarize(
+    Min = min(end_date, na.rm = TRUE),
+    Max = max(end_date, na.rm = TRUE)
+  )
+
+end_date_summary %>% 
+  kable() %>%
+  kable_styling()
+
+```
+
 # Measurement
 
-The dataset comes from fivethirtyeight.com, a statistical analysis website dedicated to US sports and elections [@fivethrityeight].
+The dataset comes from fivethirtyeight.com, a statistical analysis website dedicated to US sports and elections [@fivethirtyeight].
 Fivethirtyeight takes polls conducted on a particular topic (in this case the 2024 US presidency), and aggregates them in one dataset.
 
-Election polls are used to gauge pulic opinion about candidates, political topics, voter angagement, and public opinion of the election process.
-Election polls also serve as key insight to political organisations representing candidates, polls can show potenital weak points are low support in areas and populations.
+Election polls are used to gauge public opinion about candidates, political topics, voter engagement, and public opinion of the election process.
+Election polls also serve as key insight to political organisations representing candidates, polls can show potential weak points across demographics.
 
@@
 The key distinction is that election polls typically try to target *voters*, or people who are most likely to vote in the election cycle being investigated.
-This is a key challenge to pollsters, as there are many subgroups of voters, such as people who seldom vote or almost never vote.
-According to Pew Research, a third of US citizens did not vote in the 2020 election, despite that election cycle having high voter turnout [@pew_polls].
-There may also be cases when repondents of a poll saying they will/will not vote change their mind come election day.
+This is a key challenge to pollsters, as there are many subgroups of voters, such as people who seldom vote or never vote.
+According to Pew Research, a third of US citizens did not vote in the 2020 election, despite the 2020 election cycle having high voter turnout [@pew_polls].
+There may also be cases when respondents of a poll reconsider whether they will vote come election day.
 
 Election polling has exploded in recent times according to Pew Research.
-They suggest that double thre are double the amount of active pollsters in 2022 compared to 2000 [@pew_polls].
-One reason mentioned is the advent of online polling, which is simple, cheap, and can easily reach US citizens across the country.
-As a result, while the diversity of pollsters has increased dramatically, this also means there are as many amateur polls being conducted.
+They suggest that double there are double the amount of active pollsters in 2022 compared to 2000 [@pew_polls].
+One reason mentioned is the advent of online polling, which is simple, cheap, and can easily reach US citizens across the country, compared to traditional methods such as poll by phone and mail.
+As a result, while the diversity of pollsters has increased dramatically, the variance in results methodology, and ethics have also increased
 
@@
 Between the 2016 and 2020 elections, there has been a trend of pollsters not fully capturing Trump's support.
-For example in 2020, despite Biden winning the election, there were reports from the American Associcate for Public Opinion Reseach that national polls overstated his victory over Trump by 3.9 percentage points [@pew_polls].
+For example in 2020, despite Biden winning the election, there were reports from the American Associcate for Public Opinion Research that national polls overstated his victory over Trump by 3.9 percentage points [@pew_polls].
 
-Fivethirty eight mitigates the influx of pollsters and the potential for errors by routinely checking up on pollsters as they produce reports, wther they have an established track record, or if they are just new to data presentation, There are 2 types of checks fivethirtyeight uses: Methodology and Ethics.
+Fivethirty eight mitigates the influx of pollsters and the potential for errors by routinely checking up on pollsters as they produce reports.
+Fivethirtyeight checks every pollster, regardless of experience or prestige.
+There are 2 types of checks fivethirtyeight uses: Methodology and Ethics.
 
@@
 If unclear, fivethirtyeight will do a more thorough investigation of the pollster, and ask the organizers questions about their methodology and ethics if none were visible in the public report.
-If there are any violations in ethics, the severity on consequences will vary on wjat the offence is.
-If there is evidence of falifyin data or engaging in betting, the pollster is blacklisted from their data, any polls that come out of this pollster are extremely dubious and not trustworthy to build a model or report on.
+If there are any violations in ethics, the severity on consequences will vary on what the offence is.
+If there is evidence of falsifying data or engaging in betting markets, the pollster is blacklisted from their data.
 Otherwise, pollsters are given a chance to fix ethical issues presented by fivethirtyeight.
-Failiure to do so results in the pollster and any of its polls being taken off the dataset, but if the pollster shows it has updated its ethics policies and applied them to recent reports, they may be allowed back in [@faq538].
+Failure to do so results in the pollster and any of its polls being taken off the dataset, but fivethirtyeight can reverse this decision if the pollster demonstrates satifactory ethics and methodology practices[@faq538].
 

### 2024-11-04 scripts/03-clean_data.R
Updated clean_data script to Prankit's version, rename raw data and analysis data

@@ -16,16 +16,18 @@ library(rstanarm)
 library(splines)
 library(arrow)
 
+#### Workspace setup ####
+library(tidyverse)
+
 #### Clean data ####
-# Read in the data and clean variable names
-data <- read_csv("data/01-raw_data/president_polls.csv") |>
+data <- read_csv("data/01-raw_data/raw_elections_data.csv") |>
   clean_names()
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 3 # Need to investigate this choice - come back and fix. 
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
     # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
   ) |>
   mutate(
@@ -37,8 +39,21 @@ just_harris_high_quality <- data |>
     num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
+# Filter data to Trump estimates based on high-quality polls after he declared
+just_trump_high_quality <- data |>
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.7
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # Update with Trump's declaration date if needed
+  mutate(
+    num_trump = round((pct / 100) * sample_size, 0)
+  )
+
 #### Save data ####
-write_csv(just_harris_high_quality, "data/02-analysis_data/harris.csv")
-
-write_parquet(just_harris_high_quality, "data/02-analysis_data/harris.parquet")
-
+write_csv(just_harris_high_quality, "data/02-analysis_data/harris_data.csv")
+write_csv(just_trump_high_quality, "data/02-analysis_data/trump_data.csv")

### 2024-11-04 paper/paper.qmd
Updated Data section, included figures and tables, added introductory paragraphs to Data section

@@
 library(rstanarm)
+library(dplyr)
 
+# Read in data from clean_data.r
 just_harris_high_quality <-
-  read_csv(file = here::here("data/02-analysis_data/harris.csv"))
+  read_csv(file = here::here("data/02-analysis_data/harris_data.csv"))
+
+just_trump_high_quality <-
+  read_csv(file = here::here("data/02-analysis_data/trump_data.csv"))
 
@@
 Each observation is a poll conducted on the 2024 US Presidential Election.
-Pollsters are required by fivethirtyeight to disclose certain information about their poll, either publicly, or through an investigation.
-Of this information, a handful are key to the predictive model.
+Fivethirtyeight's dataset was chosen due to it's comprehensive review of polls and pollsters, attributing grades in error and bias, as well as in transparency.
+Their thorough investigations ensure their dataset not only contains insightful polls, but also includes as much relevant information about the polls conducted as possible.
+This gives us plenty of predictors when building a model that will caluclate the percentage win of either Trump or Harris.
+
+The dataset was read into R [@citeR], and cleaned using the R package tidyverse [@tidyverse].
+Due to Kamal's late entry into the race, some polls do not have any data on her specifically.
+This is why we set a cut-off date at 2024/07/21, the day Kamala replaced Joe Biden as the Democratic nominee.
+
+Fivethirtyeight includes a metric called `numeric_grade`, which combines the pollscore and transparency score.
+We only included polls with a `numeric_grade` at or above 2.7.
+This gave us the best balance between quantity and quality.
+Finally, we create 2 datasets, one with polls predicting a Harris win, the other a Trump win.
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+#### Create end_date - pct base plots for Trump and Harris ####
+base_plot_kamala <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+base_plot_trump <- ggplot(just_trump_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Trump percent", x = "Date")
+```
 
@@
 ```{r}
-#| include: true
+#| label: fig-pollscore_harris
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Harris' win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their pollscore, which is how fivethirtyeight measures error and bias in polls.
+The lower the pollscore, the better the poll.
+#| echo: false
 #| warning: false
@@
 
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
-
 # Color by pollscore
-base_plot +
+base_plot_kamala +
   geom_point(aes(color = factor(pollscore))) +
@@
   theme(legend.position = "bottom")
+```
 
-pollscore_summary <- just_harris_high_quality %>%
+```{r}
+#| label: tbl-pollscore_harris
+#| tbl-cap: Summary of pollscores amonst polls predicting a Kamala Harris win. The lower the pollscore, the less prone the polls are to errors and bias. 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generating summary statistics for Harris
+pollscore_harris_summary <- just_harris_high_quality %>%
   summarize(
@@
 
-pollscore_summary %>% 
+# Displaying summary
+pollscore_harris_summary %>% 
+  kable() %>%
+  kable_styling()
+```
+
+```{r}
+#| label: fig-pollscore_trump
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Trump's win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their pollscore, which is how fivethirtyeight measures error and bias in polls.
+The lower the pollscore, the better the poll.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Color by pollscore
+base_plot_trump +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: tbl-pollscore_trump
+#| tbl-cap: Summary of pollscores amonst polls predicting a Donald Trump win. The lower the pollscore, the less prone the polls are to errors and bias. 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generating summary statistics for Harris
+pollscore_trump_summary <- just_trump_high_quality %>%
+  summarize(
+    Min = min(pollscore, na.rm = TRUE),
+    Mean = mean(pollscore, na.rm = TRUE),
+    Median = median(pollscore, na.rm = TRUE),
+    Max = max(pollscore, na.rm = TRUE)
+  )
+
+# Display summary
+pollscore_trump_summary %>% 
   kable() %>%
@@
 ```{r}
-#| include: true
+#| label: fig-samplesize_harris
+#| fig-cap: Polls require input from a subset of the target popoulation. This gpah shows the most common sizes for these subsets in polls that predicted a Harris win.
+#| echo: false
 #| warning: false
@@
 #### Plot data ####
-base_plot +
-  geom_point(aes(color = factor(sample_size))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
+just_harris_high_quality %>%
+  ggplot(mapping=aes(x = sample_size)) +
+  geom_histogram()
+```
 
+```{r}
+#| label: fig-samplesize_trump
+#| fig-cap: Polls require input from a subset of the target popoulation. This gpah shows the most common sizes for these subsets in polls that predicted a Trump win.
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot data ####
+just_trump_high_quality %>%
+  ggplot(mapping=aes(x = sample_size)) +
+  geom_histogram()
 ```
@@
 just_harris_high_quality %>%
-  ggplot(mapping=aes(x = state), y=count(n)) +
+  ggplot(mapping=aes(y = state), x=count(n)) +
   geom_bar()
 
-state_summary <- just_harris_high_quality %>%
-  summarize(
-    Min = min(count(state), na.rm = TRUE),
-    Max = max(state, na.rm = TRUE)
-  )
-
-state_summary %>% 
-  kable() %>%
-  kable_styling()
-
+just_trump_high_quality %>%
+  ggplot(mapping=aes(y = state), x=count(n)) +
+  geom_bar()
 ```

### 2024-11-04 paper/paper.qmd
Edited data, added all graphs and tables

--- 2024-11-04 paper/paper.qmd
@@
 library(rstanarm)
+library(dplyr)
 
+# Read in data from clean_data.r
 just_harris_high_quality <-
-  read_csv(file = here::here("data/02-analysis_data/harris.csv"))
+  read_csv(file = here::here("data/02-analysis_data/analysis_data_harris.csv"))
 
-bayesian_model <-
-  readRDS(file = here::here("models/electionmodel.rds"))
-
-spline_model <- 
-  readRDS(file = here::here("models/splinemodel.rds"))
+just_trump_high_quality <-
+  read_csv(file = here::here("data/02-analysis_data/analysis_data_trump.csv"))
 ```
@@
 Each observation is a poll conducted on the 2024 US Presidential Election.
-Pollsters are required by fivethirtyeight to disclose certain information about their poll, either publicly, or through an investigation.
-Of this information, a handful are key to the predictive model.
+Fivethirtyeight's dataset was chosen due to it's comprehensive review of polls and pollsters, attributing grades in error and bias, as well as in transparency.
+Their thorough investigations ensure their dataset not only contains insightful polls, but also includes as much relevant information about the polls conducted as possible.
+This gives us plenty of predictors when building a model that will caluclate the percentage win of either Trump or Harris.
+
+The dataset was read into R [@citeR], and cleaned using the R package tidyverse [@tidyverse].
+Due to Kamal's late entry into the race, some polls do not have any data on her specifically.
+This is why we set a cut-off date at 2024/07/21, the day Kamala replaced Joe Biden as the Democratic nominee.
+
+Fivethirtyeight includes a metric called `numeric_grade`, which combines the transparency_score and transparency score.
+We only included polls with a `numeric_grade` at or above 2.7.
+This gave us the best balance between quantity and quality.
+Finally, we create 2 datasets, one with polls predicting a Harris win, the other a Trump win.
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+#### Create end_date - pct base plots for Trump and Harris ####
+base_plot_kamala <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+base_plot_trump <- ggplot(just_trump_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Trump percent", x = "Date")
+```
 
@@
 
-### **`poll_score:`**
+### **`transparency_score`**
 
-The error and bias associated with a pollster.
-Positive values indicate more bias and errors, negative values indicate less.
-This value will tell us if biases in polls influences the predicted percentages of each candidate
+Fivethirtyeight awards a score out of 10 to each pollster based on how informative their methodology is.
+Understanding the methodology of a pollster assits in understanding any errors or biases present in their findings.
 
 ```{r}
-#| include: true
+#| label: fig-transscore_harris
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Harris' win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their transparency_score, which is fivethirtyeight's metric for transpareny in the pollster's methodology.
+#| echo: false
 #| warning: false
@@
 
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
+# Color by transparency_score
+base_plot_kamala +
+  geom_point(aes(color = factor(transparency_score))) +
+  geom_smooth(color = "blue") +
+  theme(legend.position = "bottom")
+```
 
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
+```{r}
+#| label: tbl-transscore_harris
+#| tbl-cap: Summary of transparency_scores amongst polls predicting a Kamala Harris win. The closer the values are to 10, the more trustworthy the predictions. 
+#| echo: false
+#| warning: false
+#| message: false
 
-pollscore_summary <- just_harris_high_quality %>%
+# Generating summary statistics for Harris
+transparency_score_harris_summary <- just_harris_high_quality %>%
   summarize(
-    Min = min(pollscore, na.rm = TRUE),
-    Mean = mean(pollscore, na.rm = TRUE),
-    Median = median(pollscore, na.rm = TRUE),
-    Max = max(pollscore, na.rm = TRUE)
+    Min = min(transparency_score, na.rm = TRUE),
+    Mean = mean(transparency_score, na.rm = TRUE),
+    Median = median(transparency_score, na.rm = TRUE),
+    Max = max(transparency_score, na.rm = TRUE)
   )
 
-pollscore_summary %>% 
+# Displaying summary
+transparency_score_harris_summary %>% 
+  kable() %>%
+  kable_styling()
+```
+
+```{r}
+#| label: fig-transscore_trump
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Trump's win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their transparency_score, which is fivethirtyeight's metric for transpareny in the pollster's methodology.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Color by transparency_score
+base_plot_trump +
+  geom_point(aes(color = factor(transparency_score))) +
+  geom_smooth(color = "red") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: tbl-transscore_trump
+#| tbl-cap: Summary of transparency_scores amongst polls predicting a Donald Trump win. The closer the values are to 10, the more trustworthy the predictions. 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generating summary statistics for Trump
+numeric_grade_trump_summary <- just_trump_high_quality %>%
+  summarize(
+    Min = min(transparency_score, na.rm = TRUE),
+    Mean = mean(transparency_score, na.rm = TRUE),
+    Median = median(transparency_score, na.rm = TRUE),
+    Max = max(transparency_score, na.rm = TRUE)
+  )
+
+# Display summary
+transparency_score_trump_summary %>% 
   kable() %>%
@@
 
-### `sample_size`
+### **`numeric_grade`**
 
-Polls are conducted over a subset of the targeted population (Hispanics, young people, Texans, LGBTQ+, rural communities, etc.).
+A cumulative score given to pollster that combines fivethirtyeight's `pollscore` and `transparency_score`.
+Like `transparency_score`, we believe this metric can give us insight into potential biases in each pollster and poll, which is why it is included in our model.
 
 ```{r}
-#| include: true
+#| label: fig-numgrade_harris
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Harris' win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their numeric_grade, which is fivethirtyeight's combined metric on transparency and bias/errors.
+#| echo: false
 #| warning: false
@@
 
-#### Plot data ####
-base_plot +
-  geom_point(aes(color = factor(sample_size))) +
-  geom_smooth() +
+# Color by transparency_score
+base_plot_kamala +
+  geom_point(aes(color = factor(numeric_grade))) +
+  geom_smooth(color = "blue") +
   theme(legend.position = "bottom")
-
 ```
-
-### `state`
-
-The state that each poll as conducted in. If there are any empty values, we assume these are national polls.
-
 ```{r}
-#| include: true
+#| label: tbl-numgrade_harris
+#| tbl-cap: Summary of numeric_grades amongst polls predicting a Kamala Harris win.
+#| echo: false
 #| warning: false
@@
 
-#### Plot data ####
-just_harris_high_quality %>%
-  ggplot(mapping=aes(x = state), y=count(n)) +
-  geom_bar()
-
-state_summary <- just_harris_high_quality %>%
+# Generating summary statistics for Harris
+numeric_grade_harris_summary <- just_harris_high_quality %>%
   summarize(
-    Min = min(count(state), na.rm = TRUE),
-    Max = max(state, na.rm = TRUE)
+    Min = min(numeric_grade, na.rm = TRUE),
+    Mean = mean(numeric_grade, na.rm = TRUE),
+    Median = median(numeric_grade, na.rm = TRUE),
+    Max = max(numeric_grade, na.rm = TRUE)
   )
 
-state_summary %>% 
+# Display summary
+numeric_grade_harris_summary %>% 
   kable() %>%
@@
 ```{r}
-#| include: true
+#| label: fig-numgrade_trump
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Trumps' win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on their numeric_grade, which is fivethirtyeight's combined metric on transparency and bias/errors.
+#| echo: false
 #| warning: false
@@
 
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
+# Color by transparency_score
+base_plot_kamala +
+  geom_point(aes(color = factor(numeric_grade))) +
+  geom_smooth(color = "red") +
+  theme(legend.position = "bottom")
+```
+```{r}
+#| label: tbl-transscore_trump
+#| tbl-cap: Summary of transparency_scores amongst polls predicting a Donald Trump win. The closer the values are to 10, the more trustworthy the predictions. 
+#| echo: false
+#| warning: false
+#| message: false
 
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
+# Generating summary statistics for Trump
+numeric_grade_trump_summary <- just_trump_high_quality %>%
+  summarize(
+    Min = min(transparency_score, na.rm = TRUE),
+    Mean = mean(transparency_score, na.rm = TRUE),
+    Median = median(transparency_score, na.rm = TRUE),
+    Max = max(transparency_score, na.rm = TRUE)
+  )
 
-# Color by pollster
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = pollster)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-# Facet by pollster
-# Make the line 21 issue obvious
-# Also - is there duplication???? Need to go back and check
-base_plot +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(vars(pollster))
-
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
+# Display summary
+numeric_grade_trump_summary %>% 
+  kable() %>%
+  kable_styling()
 
@@
 
-### `party_binary`
+### `pollster`
+
+```{r}
+#| label: fig-pollster_harris
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Harris' win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on the pollster who authored the poll.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Color by transparency_score
+base_plot_kamala +
+  geom_point(aes(color = pollster)) +
+  geom_smooth(color = "blue") +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: fig-pollster_trump
+#| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
+The polls in this graph predict Trump's win percentage, with the average in the blue line.
+The dots in the graph below represent individual polls conducted in the months leading up to the election.
+They are coloured based on the pollster who authored the poll.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Color by transparency_score
+base_plot_trump +
+  geom_point(aes(color = pollster)) +
+  geom_smooth(color = "red") +
+  theme(legend.position = "bottom")
+```
+
+### `states`
+
+```{r}
+#| label: fig-state
+#| fig-cap: Battleground states are what usually determines the winner of a presidential election.
+This figure outlines the polling outcomes of each candidate from August to November, for each of the seven battleground states this election cycle.
+#| echo: false
+#| warning: false
+#| message: false
+#| 
+# Define major swing states
+swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
+
+combined_data <- bind_rows(
+  just_harris_high_quality |> mutate(candidate = "Kamala Harris"),
+  just_trump_high_quality |> mutate(candidate = "Donald Trump")
+)
+
+# Filter data for Harris and Trump in the swing states
+swing_states_data <- combined_data |>
+  filter(state %in% swing_states)
+
+# Create a base plot for swing states data
+swing_states_plot <- ggplot(swing_states_data, aes(x = end_date, y = pct, color = candidate)) +
+  theme_minimal() +
+  labs(x = "Polling End Date", y = "Polling Percentage", title = "Polling Outcomes in Swing States") +
+  theme(legend.position = "bottom")
+
+# Facet the plot by state
+swing_states_plot +
+  geom_point(aes(shape = candidate), alpha = 0.7) +
+  geom_smooth(aes(linetype = candidate), se = FALSE) +
+  facet_wrap(vars(state)) +
+  scale_color_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red"))
+
+```
 
@@
 ```{r}
-#| include: true
+#| label: fig-enddate_harris
+#| fig-cap: Polls are conducted from a start date to an end date.
+The closer the end date is to election day, the more insightful the results are to analysts and the general public.
+This distribution shows the end date of polls predicting a Kamala Harris win.
+#| echo: false
 #| warning: false
@@
   ggplot(mapping=aes(x = end_date)) +
-  geom_histogram()
+  geom_histogram(fill = "blue", alpha = 0.7, color = "black",)
+```
 
-end_date_summary <- just_harris_high_quality %>%
-  summarize(
-    Min = min(end_date, na.rm = TRUE),
-    Max = max(end_date, na.rm = TRUE)
-  )
+```{r}
+#| label: fig-enddate_trump
+#| fig-cap: Polls are conducted from a start date to an end date.
+The closer the end date is to election day, the more insightful the results are to analysts and the general public.
+This distribution shows the end date of polls predicting a Donald Trump win.
+#| echo: false
+#| warning: false
+#| message: false
 
-end_date_summary %>% 
-  kable() %>%
-  kable_styling()
-
+#### Plot data ####
+just_trump_high_quality %>%
+  ggplot(mapping=aes(x = end_date)) +
+  geom_histogram(fill = "red", alpha = 0.7, color = "black",)
 ```

### 2024-11-04 paper/paper.qmd
Final edit of Data section

@@
 Fivethirtyeight includes a metric called `numeric_grade`, which combines the transparency_score and transparency score.
-We only included polls with a `numeric_grade` at or above 2.7.
+We only included polls with a `numeric_grade` at or above 2.5.
 This gave us the best balance between quantity and quality.
@@
 
+According to @tbl-transscore_harris and @tbl-transscore_trump, the difference between candidates is very minimal, with the minimum, maximum, and even average being the same (only in the media do we see a slight lead by Trump).
+Both @fig-transscore_harris & @fig-transscore_trump are similar as well, with most polls achieving a 9.0 transparency score.
 ```{r}
@@
 # Generating summary statistics for Trump
-numeric_grade_trump_summary <- just_trump_high_quality %>%
+transparency_score_trump_summary <- just_trump_high_quality %>%
   summarize(
@@
 
+Similar to `transparency_score`, @tbl-numgrade_harris & @tbl-numgrade_trump are identical except for the median, where Trump again holds a slight advantage.
+Similarly, @fig-numgrade_harris & @fig-numgrade_trump are almost similar, with a high number of 3's for numeric_grade.
+This makes sense, as `numeric_grade` is a combination of `transparency_score` and `pollscore`, so any patterns from the latter will apply to the former.
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 # Color by transparency_score
-base_plot_kamala +
+base_plot_trump +
   geom_point(aes(color = factor(numeric_grade))) +
@@
 ```
+
 ```{r}
-#| label: tbl-transscore_trump
+#| label: tbl-numgrade_trump
 #| tbl-cap: Summary of transparency_scores amongst polls predicting a Donald Trump win. The closer the values are to 10, the more trustworthy the predictions. 
@@
   summarize(
-    Min = min(transparency_score, na.rm = TRUE),
-    Mean = mean(transparency_score, na.rm = TRUE),
-    Median = median(transparency_score, na.rm = TRUE),
-    Max = max(transparency_score, na.rm = TRUE)
+    Min = min(numeric_grade, na.rm = TRUE),
+    Mean = mean(numeric_grade, na.rm = TRUE),
+    Median = median(numeric_grade, na.rm = TRUE),
+    Max = max(numeric_grade, na.rm = TRUE)
   )
@@
 
+Each poll is conducted by a pollster, who may make multiple polls leading up to the election. This may increase accuracy of their predictions, which is why it will be in our model.
+
+@fig_pollster-harris & @fig_pollster_trump show an interesting bump in predictions by George Mass. This may have been an error on their part however, as it is almost immediately rectified.
+
 ```{r}
@@
 #| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
-The polls in this graph predict Harris' win percentage, with the average in the blue line.
-The dots in the graph below represent individual polls conducted in the months leading up to the election.
-They are coloured based on the pollster who authored the poll.
+Each graph shows individual polls conducted in the months leading up to the election.
+They represent the change in prediciton for Kamala Harris winning the presidency.
 #| echo: false
@@
 
-# Color by transparency_score
+# Facet by pollster
 base_plot_kamala +
-  geom_point(aes(color = pollster)) +
+  geom_point() +
   geom_smooth(color = "blue") +
-  theme(legend.position = "bottom")
+  facet_wrap(vars(pollster))
 ```
@@
 #| fig-cap: As election day draws close, pollsters are wrapping up their polls and posting their predictions on who will win.
-The polls in this graph predict Trump's win percentage, with the average in the blue line.
-The dots in the graph below represent individual polls conducted in the months leading up to the election.
-They are coloured based on the pollster who authored the poll.
+Each graph shows individual polls conducted in the months leading up to the election.
+They represent the change in prediciton for Donald Trump winning the presidency.
 #| echo: false
@@
 
-# Color by transparency_score
+# Facet by pollster
 base_plot_trump +
-  geom_point(aes(color = pollster)) +
+  geom_point() +
   geom_smooth(color = "red") +
-  theme(legend.position = "bottom")
+  facet_wrap(vars(pollster))
 ```
@@
 ### `states`
+
+The states each poll was conducted in. This is not one of our predictors for our model, but it can provide insight into what the prediction will be.
+
+@fig-state has Harris win 4 t 7 battleground states, but the margins are increadibly close and all it would take is for one of them to be incorrect to reverse the prediction.
 
@@
 
+Both @fig-enddate_harris & @fig-enddate_trump show a large volume of polls ending at the beginning of November, which would be the most accurate findiings leading up to the election.
+
 ```{r}
@@
   geom_histogram(fill = "red", alpha = 0.7, color = "black",)
+```
+
+## Response
+
+### `pct`
+
+This the the predicted percentage each candidate will get on election day.
+We have assumed a two-party system for our report, as any third-party candidates would have a negligible impact on election day.
+
+@fig-pct_harris & @fig-pct_trump are centered just below 50% indicating a very close match.
+@tbl-pct_harris & @tbl-pct_trump show Harris leading in Mean and Median percentages however, ever so slightly.
+
+```{r}
+#| label: fig-pct_harris
+#| fig-cap: The distribution of predicted Harris vote percentage on election day.
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot data ####
+just_harris_high_quality %>%
+  ggplot(mapping=aes(x = pct)) +
+  geom_histogram(fill = "blue", alpha = 0.7, color = "black",)
+```
+
+```{r}
+#| label: tbl-pct_harris
+#| tbl-cap: Summary of percentage vote predictions for Kamala Harris.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generating summary statistics for Harris
+pct_harris_summary <- just_harris_high_quality %>%
+  summarize(
+    Min = min(pct, na.rm = TRUE),
+    Mean = mean(pct, na.rm = TRUE),
+    Median = median(pct, na.rm = TRUE),
+    Max = max(pct, na.rm = TRUE)
+  )
+
+# Display summary
+pct_harris_summary %>% 
+  kable() %>%
+  kable_styling()
+
+```
+
+```{r}
+#| label: fig-pct_trump
+#| fig-cap: The distribution of predicted Trump vote percentage on election day.
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot data ####
+just_trump_high_quality %>%
+  ggplot(mapping=aes(x = pct)) +
+  geom_histogram(fill = "red", alpha = 0.7, color = "black",)
+```
+
+```{r}
+#| label: tbl-pct_trump
+#| tbl-cap: Summary of percentage vote predictions for Donald Trump.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generating summary statistics for Trump
+pct_trump_summary <- just_trump_high_quality %>%
+  summarize(
+    Min = min(pct, na.rm = TRUE),
+    Mean = mean(pct, na.rm = TRUE),
+    Median = median(pct, na.rm = TRUE),
+    Max = max(pct, na.rm = TRUE)
+  )
+
+# Display summary
+pct_trump_summary %>% 
+  kable() %>%
+  kable_styling()
+
 ```
@@
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Posterior predictive checks
-pp_check(bayesian_model)
-
-# Summarize the model
-summary(bayesian_model)
-
-# Plot random effects
-plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
-
-```
-
-```{r}
-#| label: fig-1
-#| fig-cap: Street tree count of 10 most popoulus tree species in Toronto up to September 2024
-#| echo: false
-#| warning: false
-#| message: false
-
-# Summarize the model
-summary(spline_model)
-
-# Posterior predictive checks
-pp_check(spline_model)
-```
-
 \newpage
